---
title: "Summary: Criteria Set A"
author: "Matt Schumwinger, Big Lake Data LLC"
date: "June 7, 2016"
output: html_document
---

```{r setup, include=FALSE, eval=FALSE}
knitr::opts_chunk$set(eval = FALSE, include = FALSE)
source("global.R") # make sure global.R is set to pick up the full file trans.Rds
library(stringr)
library(data.table)
library(stringr)
library(lubridate)
library(dplyr)

filers <- data.table(filers, key = "FILER_ID")
trans <- data.table(trans, key = "FILER_ID")
# drop bad dates                                                                  TODO: see if these can be saved somehow
trans <- trans[DATE1_10 >= lubridate::ymd(19990101) 
                 & DATE1_10 < lubridate::ymd(20170101) ,]

# ----- helpers -----
ante5yrs <- function(date) {
  date - lubridate::years(5)
}
inCodes <- c("A","B","C","D","E","G")  # incoming contribution codes
outCodes <- c("F", "H")  # outgoing contribution codes
```

```{r load workspace}
load("data/sugarman.Rds")
```

Data profile (all years):  

- **`r formatC(nrow(trans), big.mark=",") `** transactions
- **`r formatC(nrow(filers), big.mark=",") `** filers

## Applying the "Sugarman Standard""
Criteria:  

 1) **The recipient is a county committee or party committee**  
```{r c1, message=FALSE, warning=FALSE, eval=FALSE}
# create table of just county or party committees
cmtes <- subset(filers, FILER_TYPE == "COMMITTEE" 
                  & !COMMITTEE_TYPE %in% c("1", "2", "9", "9B", "9U"))
table(cmtes$COMMITTEE_TYPE)
# subset transactions matching just filers in committees table
trans1 <- trans[cmtes]                                                            # BUG: this may create NAs in first column
trans1 <- trans1[, c(1:32), with=FALSE]
```
results in:  

- **`r formatC(nrow(cmtes), big.mark=",") `** filers classed as county or party committees
- **`r formatC(nrow(trans), big.mark=",") `** transactions matching such committees

 2) **The contribution amounted to at least two times the average annual amount of contributions received by the recipient county committee in the five years prior to the contribution** 
```{r c2, message=FALSE, warning=FALSE, eval=FALSE}
# subset trans by only incoming contribution codes
trans2 <- trans1 %>% filter(TRANSACTION_CODE %in% inCodes)

# calculate cumulative average of trans for that filer prior to date of trans
# Note: this is not in a 5 year window, but works
trans2 <- trans2 %>%
  arrange(FILER_NAME, DATE1_10) %>%
  group_by(FILER_NAME) %>%
  mutate(cumAvg = cummean(AMOUNT_70)) %>%
  mutate(bigTrans = AMOUNT_70 >= 2*cumAvg) 

tBig <- filter(trans2, bigTrans)

```
results in:  

 - **`r formatC(nrow(tBig), big.mark=",") `** transactions received by a county or party committee that were at least two times the average annual amount of contributions received by the recipient committee prior to the contribution. (NOTE: this is greater than a 5-yr window.)

 3) **The contributor had given to the recipient committee no more than twice in the past**
```{r c3, message=FALSE, warning=FALSE, eval=FALSE}
# one-off name fix
trans2[FILER_NAME == "*PHELPS DEMOCRATIC COMMITTEE", 
       FILER_NAME := "PHELPS DEMOCRATIC COMMITTEE"]

# create "oppositeParty" field
trans3 <- trans2 %>%                                                            # TODO: condition this on time before contribution 
  mutate(opp_party = paste(CORP_30, FIRST_NAME_40, MID_INIT_42, LAST_NAME_44)) 

# count contributions by opposite party
count_by_opp_part <- trans3 %>%
  group_by(FILER_NAME, opp_party) %>% 
  summarise(tot_from_opp_party = n())
trans3 <- left_join(trans3, count_by_opp_part, by = c("FILER_NAME", "opp_party")) #TODO: work on encoding warning message
trans3 <- trans3 %>% mutate(rare_opp_party = tot_from_opp_party <= 3)
# table(trans3$reg_opp_party) 
tBigRare <- filter(trans3, bigTrans & rare_opp_party)
```
results in:  

 - **`r formatC(nrow(tBigRare), big.mark=",") `** transactions received by a county or party committee that were:  
 
 - at least two times the average annual amount of contributions received by the recipient committee prior to the contribution. (NOTE: this is greater than a 5-yr window.) AND  
 - were from a contributor that had given to the recipient committee not more than twice ever else. (NOTE: this is broader than looking just in the time preceeding the contribution.)

 4) **No more than one month after receiving the contribution(s), the county committee transferred a similar amount of money directly to a candidate's campaign (i.e., not independent expenditures).**
```{r c4, message=FALSE, warning=FALSE, eval=FALSE}

df <- setDT(tBigRare)                                                           
                                                                                
# create direction column of target transactions and filter for in and out
df[, direction := ifelse(TRANSACTION_CODE %in% inCodes, "in", 
                         ifelse(TRANSACTION_CODE %in% outCodes, "out", NA))]
df <- df %>% filter(!is.na(direction)) %>%
  mutate(end_30 = DATE1_10 - days(30)) %>%
  select(FILER_NAME, DATE1_10, end_30, direction, TRANSACTION_CODE, AMOUNT_70, tID)

# create reference tables
justIns <- df %>% filter(direction == "in")                                       # this should be redundent
justOuts <- trans1 
justOuts[, direction := ifelse(TRANSACTION_CODE %in% inCodes, "in",               # TODO: refactor this to just dplyr for both df and trans1
                         ifelse(TRANSACTION_CODE %in% outCodes, "out", NA))]
justOuts <- justOuts %>% filter(!is.na(direction)) %>%
  mutate(end_30 = DATE1_10 - days(30)) %>%
  select(FILER_NAME, DATE1_10, end_30, direction, TRANSACTION_CODE, AMOUNT_70, tID)

justIns[, `:=`(DATE1_102 = DATE1_10, id = .I)]
setkey(justOuts, FILER_NAME, end_30, DATE1_10)

ans = foverlaps(justIns, justOuts, by.x=c("FILER_NAME", "DATE1_10", "DATE1_102"))[, DATE1_102 := NULL]

flagged <- ans %>% filter(!is.na(direction)) %>%
  # group_by(i.tID, FILER_NAME, TRANSACTION_CODE) %>% # this is grouped by outgoing transaction type
  group_by(i.tID, FILER_NAME) %>%
  summarise(
            nOutTrans = n(),
            totalOut = sum(AMOUNT_70, na.rm = TRUE)) %>%
  mutate(tID = i.tID)

flagged <- inner_join(df, flagged, by = "tID")

flagged <- flagged %>% select(-c(FILER_NAME.y, direction, end_30, i.tID))
names(flagged) <- c("tID","FILER_NAME", "DATE1_10","IN_TRANSACTION_CODE",
                    "AMOUNT_70","nOutTrans","totalOut" )
```
results in:  

 - **`r formatC(nrow(flagged), big.mark=",") `** incoming contributions to a county or party committee where that comittee made an outgoing contribution within 30 days. (NOTE: this is not conditioned on recipient type or "similar amount")

```{r save workspace, eval=FALSE, include=FALSE}
## ---- save workspace ---
rm(ans, df, justIns, justOuts)
save.image("data/sugarman.Rds")

path <- paste0("data/flagged_", Sys.Date(), ".csv")
write.csv(flagged, path, row.names = F)

```
