---
title: "Summary"
author: "Matt Schumwinger, Big Lake Data LLC"
date: "June 2, 2016"
output: html_document
---

```{r setup, include=FALSE, eval=FALSE}
knitr::opts_chunk$set(eval = FALSE, include = FALSE)
source("global.R") # make sure global.R is set to pick up the full file trans.Rds
library(stringr)
library(data.table)
library(stringr)
library(lubridate)
```

```{r load workspace}
load("data/sugarman.Rds")
```
Data profile:  

- **`r formatC(nrow(trans), big.mark=",") `** transactions in election years 2008, 2010, 2012, and 2014
- **`r formatC(nrow(filers), big.mark=",") `** filers

## Applying the Sugarman Standard
Criteria:  

 1) The recipient is a county committee or party committee;  
```{r c1, message=FALSE, warning=FALSE, eval=FALSE}
# create table of just county or party committees
cmtes <- subset(filers, FILER_TYPE == "COMMITTEE" 
                  & !COMMITTEE_TYPE %in% c("1", "2", "9", "9B", "9U"))
table(cmtes$COMMITTEE_TYPE)
# subset transactions matching just filers in committees table
trans1 <- trans[cmtes]
trans1 <- trans1[, c(1:31), with=FALSE]
```
results in:  
- **`r formatC(nrow(cmtes), big.mark=",") `** filers classed as county or party committees
- **`r formatC(nrow(trans), big.mark=",") `** transactions matching such committees

 2) The contribution amounted to at least two times the average annual amount of contributions received by the recipient county committee in the five years prior to the contribution;  
```{r c2, message=FALSE, warning=FALSE, eval=FALSE}
# subset trans by only incoming contribution codes
trans2 <- trans1 %>% filter(TRANSACTION_CODE %in% inCodes)
```
 3) The contributor had given to the recipient committee no more than twice in the past; and  
```{r c3, message=FALSE, warning=FALSE, eval=FALSE}

```
 4) No more than one month after receiving the contribution(s), the county committee transferred a similar amount of money directly to a candidate's campaign (i.e.not independent expenditures).
```{r c4, message=FALSE, warning=FALSE, eval=FALSE}

```


```{r munge, eval=FALSE, include=FALSE}

filers <- data.table(filers, key = "FILER_ID")
trans <- data.table(trans, key = "FILER_ID")
# drop bad dates
trans <- trans[DATE1_10 >= lubridate::ymd(19990101) 
                 & DATE1_10 < lubridate::ymd(20170101) ,]

# ----- helpers -----
ante5yrs <- function(date) {
  date - lubridate::years(5)
}
inCodes <- c("A","B","C","D","E","G")  # incoming contribution codes

# ---- criterion #1 ----
# create table of just county or party committees
cmtes <- subset(filers, FILER_TYPE == "COMMITTEE" 
                  & !COMMITTEE_TYPE %in% c("1", "2", "9", "9B", "9U"))
table(cmtes$COMMITTEE_TYPE)
# subset transactions matching just filers in committees table
trans1 <- trans[cmtes]
trans1 <- trans1[, c(1:31), with=FALSE]

# ---- criterion #2 ----

# subset trans by only incoming contribution codes
trans2 <- trans1 %>% filter(TRANSACTION_CODE %in% inCodes)

# ---- criterion #3 ----
# one-off name fix
trans2[FILER_NAME == "*PHELPS DEMOCRATIC COMMITTEE", FILER_NAME := "PHELPS DEMOCRATIC COMMITTEE"]

# create "oppositeParty" field
# TODO: condition this on time before contribution like in criterion 2
trans3 <- trans2 %>%
  mutate(opp_party = paste(CORP_30, FIRST_NAME_40, MID_INIT_42, LAST_NAME_44)) 
count_by_opp_part <- trans3 %>%
  group_by(FILER_NAME, opp_party) %>% 
  summarise(tot_from_opp_party = n())
trans3 <- left_join(trans3, count_by_opp_part, by = c("FILER_NAME", "opp_party")) #TODO: work on warning message
trans3 <- trans3 %>% mutate(reg_opp_party = tot_from_opp_party > 3)
# table(trans3$reg_opp_party)  


# ---- criterion #2 (cont.) ----

# ---- TODO: convert to lapply and move back before criterion 3
add5yrAvg <- function(dt){
  start <- Sys.time()
  for (i in 1:nrow(dt)) {
    filer <- dt$FILER_NAME[i]
    date <- dt$DATE1_10[i]
    yrs5 <- interval(ante5yrs(date), date)
    
    # append preceding 5-yr contribution average by filer
    tt <- subset(dt, DATE1_10 %within% yrs5 & FILER_NAME == filer, select = AMOUNT_70)
    avg <- mean(tt$AMOUNT_70)
    dt$pre5yrAvg[i] <- avg
    # is AMOUNT 2x more than that year's average?
    dt$bigTrans[i] <- dt$AMOUNT_70[i] > 2*avg
  }
  end <- Sys.time()
  print(end - start)
  dt
}

trans3F <- trans3 %>% filter(!reg_opp_party)
trans3F2 <- add5yrAvg(trans3F)
test <- trans3F2 %>% filter(bigTrans)

## ---- save workspace ---
save.image("data/sugarman.Rds")


```
