---
title: 'Summary: Criteria Set A'
author: "Matt Schumwinger, Big Lake Data LLC"
date: "June 8, 2016"
output: html_document
---

```{r setup, include=FALSE, eval=FALSE}
setwd("~/Documents/GitHub_projects/NYSBOE3")
knitr::opts_chunk$set(eval = FALSE, include = FALSE)
source("global.R") # make sure global.R is set to pick up the full file trans.Rds
library(stringr)
library(data.table)
library(stringr)
library(lubridate)
library(dplyr)
library(knitr)

filers <- data.table(filers, key = "FILER_ID")
trans <- data.table(trans, key = "FILER_ID")
# drop bad dates                                                                  TODO: see if these can be saved somehow
trans <- trans[DATE1_10 >= lubridate::ymd(19990101) 
                 & DATE1_10 < lubridate::ymd(20170101) ,]

nTrans <- formatC(nrow(trans), big.mark=",")
nFilers <- formatC(nrow(filers), big.mark=",")

# ----- helpers -----
ante5yrs <- function(date) {
  date - lubridate::years(5)
}
# inCodes <- c("A","B","C","D","E","G")  # incoming contribution codes
# outCodes <- c("F", "H")  # outgoing contribution codes
```

```{r load workspace, include=FALSE}
setwd("~/Documents/GitHub_projects/NYSBOE3")
load("data/sugarman.Rds")
```

### Applying the four-part "Sugarman Standard"

NYSBOE data profile (all years):  

- **`r nTrans `** transactions
- **`r nFilers `** filers
  
**Criterion #1: The recipient is a county committee or party committee.**
```{r c1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# create table of just county or party committees
cmtes <- subset(filers, FILER_TYPE == "COMMITTEE" 
                  & COMMITTEE_TYPE %in% c("3","3H", "6", "6H"))
table(cmtes$COMMITTEE_TYPE)
# subset transactions matching just filers in committees table
trans1 <- trans[cmtes] 
# prune "empty" rows added from join
trans1 <- filter(trans1, !is.na(FILER_NAME))
trans1 <- trans1[, c(1:32), with=FALSE]

nCmtes <- formatC(nrow(cmtes), big.mark=",")
tableCmtes <- data.frame(table(cmtes$COMMITTEE_TYPE))
names(tableCmtes) <- c("Committee Type", "Count")
nTrans1 <- formatC(nrow(trans1), big.mark=",") 
```

- **`r nCmtes `** "middleman" filers classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping entities (Committee Type = "3","3H", "6", "6H"). 
- **`r nTrans1`** transactions matching such filing entities.

**Criterion #2: The contribution amounted to at least two times the average annual amount of contributions received by the recipient county committee in the five years prior to the contribution.** 
```{r c2, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# subset trans by only incoming contribution codes
inCodes <- c("A","B","C","D","E","G")  # incoming contribution codes
trans2 <- trans1 %>% filter(TRANSACTION_CODE %in% inCodes)

# calculate cumulative average of trans for that filer prior to date of trans
# Note: this is not in a 5 year window, but works
trans2 <- trans2 %>%
  arrange(FILER_NAME, DATE1_10) %>%
  group_by(FILER_NAME) %>%
  mutate(cumAvg = cummean(AMOUNT_70)) %>%
  mutate(bigTrans = AMOUNT_70 >= 2*cumAvg) 

tBig <- filter(trans2, bigTrans)

nTrans2 <- formatC(nrow(tBig), big.mark=",")
```

 - **`r nTrans2 `** in-coming transactions (TRANSACTION CODE: "A","B","C","D","E","G") received by a  "middleman" filer (classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping) that were at least two times the average amount of contributions received by that "middleman" prior to the transaction. (NOTE: this is broader than a 5-year window.)

**Criterion #3: The contributor had given to the recipient committee no more than twice in the past.**
```{r c3, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# create "oppositeParty" field
trans3 <- trans2 %>%                                                            # TODO: condition this on time before contribution
  mutate(opp_party = paste(CORP_30, FIRST_NAME_40, MID_INIT_42, LAST_NAME_44)) 

# count contributions by opposite party
count_by_opp_part <- trans3 %>%   # TODO: inlcude and document this table in app
  group_by(FILER_NAME, opp_party) %>% 
  summarise(tot_from_opp_party = n())
trans3 <- left_join(trans3, count_by_opp_part, by = c("FILER_NAME", "opp_party")) 
trans3 <- trans3 %>% mutate(rare_opp_party = tot_from_opp_party <= 3)

tBigRare <- filter(trans3, bigTrans & rare_opp_party)

ntBigRare <- formatC(nrow(tBigRare), big.mark=",")
```

 - **`r ntBigRare `** transactions received by a "middleman" filer (classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping) that were:  
    - at least two times the average amount of contributions received by the "middleman" prior to the contribution. (NOTE: this is greater than a 5-yr window.) ***AND***  
    - from a contributor that had given to the "middleman" not more than twice ever else. (NOTE: this is broader than looking just in the time preceding the contribution.)

**Criterion #4: No more than one month after receiving the contribution(s), the county committee transferred a similar amount of money directly to a candidate's campaign (i.e., not independent expenditures).**
```{r c4, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

df <- setDT(tBigRare)                                                           
                                                                                
# create direction column of target transactions and filter for in and out
df[, direction := ifelse(TRANSACTION_CODE %in% inCodes, "in", 
                         ifelse(TRANSACTION_CODE %in% outCodes, "out", NA))]
df <- df %>% filter(!is.na(direction)) %>%
  mutate(end_30 = DATE1_10 - days(30)) %>%
  select(FILER_NAME, DATE1_10, end_30, direction, TRANSACTION_CODE, AMOUNT_70, tID)

# create temp tables
justIns <- df  # %>% filter(direction == "in") # this is redundent

justOuts <- trans1 
outCodes <- c("F", "H")  # outgoing contribution codes
justOuts[, direction := ifelse(TRANSACTION_CODE %in% inCodes, "in",               # TODO: refactor this to just dplyr for both df and trans1
                         ifelse(TRANSACTION_CODE %in% outCodes, "out", NA))]
justOuts <- justOuts %>% filter(direction == "out") %>% 
  mutate(end_30 = DATE1_10 - days(30)) %>%
  select(FILER_NAME, DATE1_10, end_30, direction, TRANSACTION_CODE, AMOUNT_70, tID)

justIns[, `:=`(DATE1_102 = DATE1_10, id = .I)]
setkey(justOuts, FILER_NAME, end_30, DATE1_10)

ans = foverlaps(justIns, justOuts, by.x=c("FILER_NAME", "DATE1_10", "DATE1_102"))[, DATE1_102 := NULL]

flagged <- ans %>% 
  filter(!is.na(direction)) %>% # filters zero rows out
  # group_by(i.tID, FILER_NAME, TRANSACTION_CODE) %>% # this is grouped by outgoing transaction type
  group_by(i.tID, FILER_NAME) %>%
  summarise(nOutTrans = n(),
            totalOut = sum(AMOUNT_70, na.rm = TRUE)) %>%
  mutate(tID = i.tID)

flagged <- inner_join(df, flagged, by = "tID")

flagged <- flagged %>% select(-c(FILER_NAME.y, direction, end_30, i.tID, DATE1_102, id))
names(flagged) <- c("tID","FILER_NAME", "DATE1_10","IN_TRANSACTION_CODE",
                    "AMOUNT_70","nOutTrans","totalOut" )

nFlagged <- formatC(nrow(flagged), big.mark=",")
flaggedA <- flagged
```

 - **`r nFlagged `** incoming contributions (TRANSACTION CODE: "A","B","C","D","E","G") to a "middleman" filer (classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping) where that "middleman" made an outgoing contribution (TRANSACTION CODE: "F", "H") within 30 days. (NOTE: this is not conditioned on recipient type (i.e., "candidates campaign") or "similar amount.")

```{r save workspace, eval=FALSE, include=FALSE}
## ---- save workspace ---
rm(ans, df, justIns, justOuts, cmtes, trans, trans1, trans2, trans3, tBig, tBigRare, flagged)
save.image("data/sugarman.Rds")
saveRDS(setDT(flaggedA), "data/flaggedA.Rds")

path <- paste0("data/flagged_", Sys.Date(), ".csv")
write.csv(flagged, path, row.names = F)

```
