---
title: 'Summary: Criteria Set B'
author: "Matt Schumwinger, Big Lake Data LLC"
date: "June 8, 2016"
output: pdf_document
---

```{r setup, include=FALSE, eval=FALSE}
setwd("~/Documents/GitHub_projects/NYSBOE3")
knitr::opts_chunk$set(eval = FALSE, include = FALSE)
source("global.R") # make sure global.R is set to pick up the full file trans.Rds
library(stringr)
library(data.table)
library(stringr)
library(lubridate)
library(dplyr)
library(knitr)

filers <- data.table(filers, key = "FILER_ID")
trans <- data.table(trans, key = "FILER_ID")
# drop bad dates                                                                  TODO: see if these can be saved somehow
trans <- trans[DATE1_10 >= lubridate::ymd(19990101) 
                 & DATE1_10 < lubridate::ymd(20170101) ,]

nTrans <- formatC(nrow(trans), big.mark=",")
nFilers <- formatC(nrow(filers), big.mark=",")

# ----- helpers -----
ante5yrs <- function(date) {
  date - lubridate::years(5)
}
# inCodes <- c("A","B","C","D","E","G")  # incoming contribution codes
# outCodes <- c("F", "H")  # outgoing contribution codes
```

```{r load workspace, include=FALSE}
setwd("~/Documents/GitHub_projects/NYSBOE3")
load("data/bird.Rds")
```

### Applying Bird's steps from June 7th email
1. Do a VLOOKUP to find the Committee Type for all Filer IDs in a given dataset.
2. Filter transactions attributed to Type 3 committees.
3. Filter "TRANSACTION_CODE" for Schedule H transactions (i.e. transfers out).
4. Filter "AMOUNT_70" for amounts greater than or equal to $25,000
5. Filter "DATE1_10" for dates in September, October or November (i.e. just before an election)

NYSBOE data profile (all years):  

- **`r nTrans `** transactions
- **`r nFilers `** filers
  
**Steps 1-2: Filter transactions attributed to Type 3 committees.**
```{r c1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# create table of just county or party committees
cmtes <- subset(filers, FILER_TYPE == "COMMITTEE" 
                  & COMMITTEE_TYPE %in% c("3","3H", "6", "6H"))
table(cmtes$COMMITTEE_TYPE)
# subset transactions matching just filers in committees table
trans1 <- trans[cmtes] 
# prune "empty" rows added from join
trans1 <- filter(trans1, !is.na(FILER_NAME))
trans1 <- trans1[, c(1:32), with=FALSE]

nCmtes <- formatC(nrow(cmtes), big.mark=",")
tableCmtes <- data.frame(table(cmtes$COMMITTEE_TYPE))
names(tableCmtes) <- c("Committee Type", "Count")
nTrans1 <- formatC(nrow(trans1), big.mark=",") 
```

- **`r nCmtes `** "middleman" filers classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping entities (Committee Type = "3","3H", "6", "6H"). 
- **`r nTrans1`** transactions matching such filing entities.

**Step 3: Filter "TRANSACTION_CODE" for Schedule H transactions (i.e. transfers out).** 
```{r c2, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# subset trans by only incoming contribution codes
inCodes <- c("A","B","C","D","E","G")  # incoming contribution codes
outCodes <- c("H")  # outgoing contribution codes
trans2 <- trans1 %>% filter(TRANSACTION_CODE %in% outCodes)

nTrans2 <- formatC(nrow(trans2), big.mark=",")
```

 - **`r nTrans2 `** out-going transactions (TRANSACTION CODE: "H") made by a "middleman" filer (classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping).

**Step 4: Filter "AMOUNT_70" for amounts greater than or equal to $25,000**
```{r c3, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# create "oppositeParty" field
trans3 <- trans2 %>%
  filter(AMOUNT_70 >= 25000)

nTrans3 <- formatC(nrow(trans3), big.mark=",")
```

 - **`r nTrans3 `** out-going transactions (TRANSACTION CODE: "H") made by a "middleman" filer (classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping) **THAT** are greater or equal to $25,000.

**Step 5: Filter "DATE1_10" for dates in September, October or November (i.e. just before an election)**
```{r c4, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
flaggedB <- trans3 %>%
  filter(lubridate::month(trans3$DATE1_10) %in% c(9,10,11))

nFlagged <- formatC(nrow(flaggedB), big.mark=",")
```

 - **`r nFlagged `** out-going transactions (TRANSACTION CODE: "H") made by a "middleman" filer (classed as CONSTITUTED COUNTY or PARTY STATE or associated housekeeping) **THAT** are greater or equal to $25,000 **AND** occurred in the months of September, October, or November.

```{r save workspace, eval=FALSE, include=FALSE}
## ---- save workspace ---
rm(cmtes, trans, trans1, trans2, trans3, flaggedA)
save.image("data/bird.Rds")
saveRDS(setDT(flaggedB), "data/flaggedB.Rds")

path <- paste0("data/flaggedB_", Sys.Date(), ".csv")
write.csv(flaggedB, path, row.names = F)

```
